---
title: "p8105_hw1_jl7035"
author: "Jeffrey Lin"
date: "2024-09-17"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load necessary libraries
```{r}
library(dplyr)
```


# load penguins data set
```{r}
data("penguins", package = "palmerpenguins")

```



# Exploring the Penguins Data Set
```{r}
head(penguins)

ncols_penguins <- ncol(penguins)
nrows_penguins <- nrow(penguins)

colnames_penguins <- colnames(penguins)
types_of_species <- unique(penguins %>%
                  select("species"))
which_islands <- unique(penguins %>%
                        select("island"))

mean_flipper_length <- penguins %>% 
                        summarise(mean_flipper_length = 
                                    mean(flipper_length_mm, na.rm = TRUE))

```

This dataset explores various characteristics regarding 3 different species
of penguin. The 3 species are as follows: Adelie, Gentoo, Chinstrap. These 
penguins come from 3 different islands: Torgersen, Biscoe, Dream. In total, 
there are 344 rows in the dataset, corresponding to 344 sampled penguins. 
There are also 8 columns in this dataset, reflecting various measures regarding 
the penguins characteristics. They are as follows: (1) "species", (2) "island",
(3) "bill_length_mm", (4) "bill_depth_mm", (5) "flipper_length_mm",
(6) "body_mass_g, (7) "sex",(8) "year. The mean length of the flipper for 
penguins in this dataset is 201 mm. 