---
title: "p8105_hw1_jl7035"
author: "Jeffrey Lin"
date: "2024-09-17"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Necessary Libraries
```{r}
library(tidyverse)
```


# Problem 1

## Load Penguins dataset
```{r}
data("penguins", package = "palmerpenguins")

```

## Describe the Penguins dataset

This dataset explores various characteristics regarding 3 different species
of penguin. The 3 species are as follows: `r unique(pull(penguins, species))`. 
These penguins come from 3 different islands: `r unique(pull(penguins, island))`
. In total, there are `r nrow(penguins)` rows and `r ncol(penguins)` columns in 
this dataset. The columns reflect various characteristics of the penguins, which 
are as follows: `r colnames(penguins)`. The mean length of the flipper for 
penguins in this dataset is `r mean(pull(penguins, flipper_length_mm), na.rm= TRUE)`. 


## Makes Scatterplot of flipper length by bill length 

```{r}
flipper_by_bill_len <- ggplot(penguins, 
                        aes(x = bill_length_mm, y = flipper_length_mm, 
                        color = species)) + 
                        geom_point()

ggsave("flipper_by_bill_len_plot.pdf")
```


# Problem 2


## Create Data Frame
```{r}
df <- tibble(
  random_sample = rnorm(10),
  greater_than_0 = (random_sample > 0),
  char_vec = c('a', 'b','c', 'd', 'e', 'f', 'g', 'h', 'i', 'j'),
  factor_vec = factor(c("small", "small", "small", "large", "large", "large",
                        "medium","medium", "medium", "medium"))
)

```

## Attempt to calculate means for each column in the table
```{r}
mean(pull(df, random_sample))
mean(pull(df, greater_than_0))
mean(pull(df,char_vec))
mean(pull(df, factor_vec))


```
Attempting to take the mean of each of the variables within df, we find that 
the mean cannot be calculated for the vector of characters or vector of factors.
This is because they are not of numeric or logical type.

## Attempt to convert each of the non-numeric columns to numeric
```{r results = 'hide', warning=FALSE}

as.numeric(pull(df, greater_than_0))
as.numeric(pull(df,char_vec))
as.numeric(pull(df, factor_vec))
            
```
Using the as.numeric function, we see that the boolean vector is converted into
numeric encodings that correspond to 0 or 1, while the factor vector 
takes the 3 levels and assigns them to corresponding integers. In both cases, 
integers are being used as categories to map the original contents of the 
vectors to respectively. Additionaly, we see that we are unable to convert 
the char vector into a numeric type, explaining why the mean could not be 
taken for this variable, as there was no corresponding numeric equivalent. This
logic also shows why the boolean vec could have its mean taken. Interestingly,
the factor variable can be converted to numeric, yet we were unable to get its
mean. Its likely that the boolean variable differs from the behavior of the 
factor vec, as the "TRUE" and "FALSE" values initially assigned to it are 
likely 0 and 1 under the hood.
